Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-16 doc/invt.tex
small changes

\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

\subsection{\label{sec:background}
Background}

\subsubsection{\label{sec:FDS}
Flat-distribution sampling}

@@

As in the single-bin case,
by varying the schedule, $\alpha(t)$, under a fixed value of
$q(T) = \int_0^T \alpha(t) \, dt$,
we can focus on the asymptotic error,
$\Err_A(T)$.
%
The value of $q(T)$ will be adjusted later
to minimize the total error in Sec. \ref{sec:optinitalpha}).
%
We can rewrite Eq. \eqref{eq:error_asym} much like an action
of a particle whose position is given by $q(t)$:
%
\begin{equation}
  \Err_A(T)
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
%\notag
\label{eq:error_asym_Lagrangian}
\end{equation}
%
where the Lagrangian is
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 1 }^{n - 1}
    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
  %\notag
  %\\
  %&=
  =
  %\sum_{ k = 1 }^{n - 1}
  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
  %\; \dot q^2( t )
  %\notag
  %\\
  %&=
  M^2\bigl(q(T) - q(t) \bigr)
  \; \dot q^2( t )
  .
\notag
\end{align}
%
where we have
used Eq. \eqref{eq:uk_def} and
defined the (square-root) mass function as
%
\begin{equation}
  M(Q)
  \equiv
  \sqrt{
    \textstyle\sum_{ k = 1 }^{n - 1}\Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
  }
  .
  \notag
  %\label{eq:mass_func}
\end{equation}
%
In this mechanical analogy,
the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
%
and the above Lagrangian, containing only the kinetic energy,
describes a free particle
with a position-dependent mass.
%
Besides, the Hamiltonian
coincides with the Lagrangian:
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L
  .
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved,
which means that the asymptotic error grows
linearly with time at a rate of $\mathcal L$,
and we may set
%
\begin{equation}
  \sqrt{ \mathcal H }
  =
  \sqrt{ \mathcal L }
  =
  M\bigl( q(T) - q(t) \bigr)
  \;
  \dot q(t)
  =
  \frac{C}{T}
  ,
  \label{eq:Lagrangian_const}
\end{equation}
%
for some positive $C$.
%
This serves as a generalization of Eq. \eqref{eq:dydt_const}.
%
By integrating this equation of motion, we get
%
\begin{equation}
  \int_{ q(T) - q(t) }^{ q(T) }
    M(Q)
    \;
    d Q
  =
  C \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
%
and $C$ can be determined from
the $t = T$ case as
%
\begin{equation}
  C =
  \int_{ 0 }^{ q(T) }
    M( Q )
    \;
    d Q
  .
  \label{eq:mint}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
The minimal asymptotic error is
%
\begin{equation}
  \Err_A(T)
  =
  \mathcal L \, T
  =
  \frac { C^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(Q) \, dQ
  \right)^2
  .
\label{eq:error_asym2}
\end{equation}

@@

We can characterize the optimal schedule
geometrically as follows.
%
First, we define
a normalized mass distribution as
%
\begin{equation}
  m(Q)
  =
  \frac{
    M(Q)
  }
  {
    \int_0^{ q(T) } M(Q') \, d Q'
  }
  =
  \frac{
    M(Q)
  }
  {
    C
  }
  ,
\notag
%\label{eq:mass_distr}
\end{equation}
%
such that
$\int_0^{q(T)} m(Q) \, dQ = 1$.
%
Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
we find that for $Q = q(T) - q(t)$,
%
\begin{align}
  m(Q)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
\label{eq:mQ_invTa}
  \\
  \int_Q^{ q(T) }
    m(Q') \, dQ'
  &=
  \frac t T
  ,
\label{eq:intmQ_tT}
\end{align}
%
which means that each point,
$\bigl(t, \alpha(t)\bigr)$,
on the schedule can be mapped
to a point,
$\bigl(Q, m(Q)\bigr)$,
on the mass distribution,
such that the ordinate of the latter curve
gives $1/(T\alpha)$,
and the area under the curve in the domain, $[Q, q(T)]$,
is equal to $t/T$,
as shown in Fig. \ref{fig:massq}.
%

@@

    \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
    \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.9);

    % xtics
    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
    \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
  \end{tikzpicture}
  %\makebox[\linewidth][c]{
  %  \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
  %}
  \caption{
    \label{fig:massq}
    Geometric characterization of the optimal schedule,
    $\alpha(t)$,
    via the normalized mass distribution, $m(Q)$.
    %
    \note{
      The figure was produced by \texttt{tikz}.
      % Gnuplot version by doc/fig/massq.gp
    }%
  }
\end{center}
\end{figure}

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  &+
  \sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  \notag \\
  &= \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(Q') \, dQ'
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(Q) \, dQ
  =
  \frac{ 2 \, C } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

%For simulations with the same $a_0$,
%the shape of the mass distribution, $m(Q)$,
%also determines how much the optimal schedule
%would depend on the simulation length, $T$.
%%
%Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
%Eq. \eqref{eq:opt_qT} shows that
%the optimal $q(T)$ increases with
%the simulation length, $T$,
%more so if $M(Q)$ has a long tail.
%%
%Thus, if $M(Q)$ decays slowly because of
%many near zero eigenvalues, $\lambda_k$'s,
%as in the metadynamics case,
%both the optimal schedule and
%normalized asymptotic error, $\Err_A(T) \cdot T$,
%would be sensitive to the simulation length, $T$.
%%
%In contrast, if all eigenvalues are $1.0$
%as in the single-bin or WL updating scheme discussed previously,
%the optimal schedule is
%given by the universal inverse-time formula,
%Eq. \eqref{eq:alpha_invt1},
%and the normalized asymptotic error
%is roughly a constant
%[cf. Eq. \eqref{eq:error_singlebin}].
%
%If $M(Q)$ decays rapidly,
%the distribution $m(Q)$ does not critically
%depend on the cutoff $q(T)$,
%once $q(T)$ is sufficiently large.
%%
%So the optimal schedules of these simulations
%approximately satisfy the scaling relations
%described in Sec. \ref{sec:mass_distr}.
%%
%If, however, $M(Q)$ has a long tail
%because of some near-zero $\lambda_k$'s,
%increasing the simulation length $T$
%generally alters the shape of the optimal schedule.

@@

Many commonly-used updating schemes,
such as the Gaussian updating scheme
used in metadynamics,
can be characterized by a rigid window
centered on the current bin.
%
For these translationally invariant or homogeneous
schemes, the eigenvalues can be readily found
from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
as a cosine transform,
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ l = 1 }^b
  \mu_l
  \cos\left(
  \frac{ k \, l \, 2 \, \pi }
       {      g \, n        }
  \right)
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
%
where $g = 1$ here for the periodic variable,
or $2$ for a non-periodic one
(cf. Appendix \ref{sec:more_wband}).

@@

  \node[draw, text width={width("Integrals, iii")}]
    (gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};

@@

  \draw[->, >=latex, thick]
    (w) edge (lambda)
    (w) edge (gamma)
    (sys) edge (gamma);

  \draw[->, >=latex, thick]
    (lambda) -| (M1)
    (gamma)  -- (M1)
    (T)      -| (M1)
    (M1)     -- (R1)
    (R1)     |- (qT);

@@

\item \label{step:gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
(cf. Appendix \ref{sec:Gamma_measure})
as well as the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it can overestimate
  $\epsilon_k$ for large-$k$ modes.
}

\item \label{step:qT}
Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
%
We implemented a variant of the Newton-Raphson method\cite{press3rd}
(for $\ln q(T)$ to avoid negative values)
with the bisection method\cite{press3rd} as a fallback.
The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
%
%Since $m(Q)$ is a decreasing function of $Q$,
%we can start from an interval $[Q_L, Q_R]$
%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
%(e.g. we can choose $Q_L = 0$,
%  and repeatedly double $Q_R$ until
%  the latter inequality is satisfied).
%%
%We can then iteratively refine the interval
%by substituting $Q_M = (Q_L + Q_R)/2$
%for either $Q_L$ and $Q_R$,
%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
%until the width of the interval is less than
%the desired precision,
%%
%and set $q(T)$ to $Q_M$.

@@

As shown in Table 1,
both the inverse-time and optimal schedule reduced
the error significantly,
and the errors roughly matched the predicted values
calculated from the precomputed $\Gamma_k$'s and $\epsilon_k$'s.
%
The Gaussian scheme appeared to be superior
to the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case.
%
In the latter case,
the error was dominated by the $k=1$ eigenmode
corresponding to the longest range motion, as $\Gamma_1 \approx \Gamma$,
and the suppression of local noise
had little effect on reducing the total error.
%
In either case, there was an optimal width of Gaussian
an overly large width could be counterproductive.
%
The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
or $0.07$ for $\varrho^* = 0.8$ in this case,
although it depended on the simulation length.

>>>>

\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{array}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

\subsection{\label{sec:background}
Flat-distribution sampling}

\subsubsection{\label{sec:FDS}
Bias potential}

@@

As in the single-bin case,
by varying the schedule, $\alpha(t)$, under a fixed value of
$q(T) = \int_0^T \alpha(t) \, dt$,
we can focus on the asymptotic error,
$\Err_A(T)$.
%
The value of $q(T)$ will be adjusted later
to minimize the total error in Sec. \ref{sec:optinitalpha}).
%
We can rewrite Eq. \eqref{eq:error_asym} much like an action
of a particle whose position is given by $q(t)$:
%
\begin{equation}
  \Err_A(T)
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
%\notag
\label{eq:error_asym_Lagrangian}
\end{equation}
%
where the Lagrangian is
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 1 }^{n - 1}
    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
  %\notag
  %\\
  %&=
  =
  %\sum_{ k = 1 }^{n - 1}
  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
  %\; \dot q^2( t )
  %\notag
  %\\
  %&=
  M^2\bigl(q(T) - q(t) \bigr)
  \; \dot q^2( t )
  .
\notag
\end{align}
%
where we have
used Eq. \eqref{eq:uk_def} and
defined the (square-root) mass function as
%
\begin{equation}
  M(q')
  \equiv
  \sqrt{
    \textstyle\sum_{ k = 1 }^{n - 1}
    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, q'}
  }
  .
  \notag
  %\label{eq:mass_func}
\end{equation}
%
In this mechanical analogy,
the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
%
and the above Lagrangian, containing only the kinetic energy,
describes a free particle
with a position-dependent mass.
%
Besides, the Hamiltonian
coincides with the Lagrangian:
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L
  .
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved,
which means that the asymptotic error grows
linearly with time at a rate of $\mathcal L$,
and we may set
%
\begin{equation}
  \sqrt{ \mathcal H }
  =
  \sqrt{ \mathcal L }
  =
  M\bigl( q(T) - q(t) \bigr)
  \;
  \dot q(t)
  =
  \frac{C}{T}
  ,
  \label{eq:Lagrangian_const}
\end{equation}
%
for some positive $C$.
%
This serves as a generalization of Eq. \eqref{eq:dydt_const}.
%
By integrating this equation of motion, we get
%
\begin{equation}
  \int_{ q(T) - q(t) }^{ q(T) }
    M(q')
    \;
    d q'
  =
  C \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
%
and $C$ can be determined from
the $t = T$ case as
%
\begin{equation}
  C =
  \int_{ 0 }^{ q(T) }
    M( q' )
    \;
    d q'
  .
  \label{eq:mint}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
The minimal asymptotic error is
%
\begin{equation}
  \Err_A(T)
  =
  \mathcal L \, T
  =
  \frac { C^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(q') \, d q'
  \right)^2
  .
\label{eq:error_asym2}
\end{equation}

@@

We can characterize the optimal schedule
geometrically as follows.
%
First, we define
a normalized mass distribution as
%
\begin{equation}
  m(\bar q)
  =
  \frac{
    M(\bar q)
  }
  {
    \int_0^{ q(T) } M(q') \, d q'
  }
  =
  \frac{
    M(\bar q)
  }
  {
    C
  }
  ,
\notag
%\label{eq:mass_distr}
\end{equation}
%
such that
$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
%
Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
we find that for $\bar q(t) \equiv q(T) - q(t)$,
%
\begin{align}
  m\left( \bar q(t) \right)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
  \label{eq:mQ_invTa}
  \\
  \int_{\bar q(t)}^{ q(T) }
    m(q') \, dq'
  &=
  \frac t T
  ,
  \label{eq:intmQ_tT}
\end{align}
%
Thus, each point on the schedule,
$\bigl(t, \alpha(t)\bigr)$,
can be mapped to a point,
$\bigl(\bar q, m(\bar q)\bigr)$,
on the mass distribution,
such that the ordinate of the latter curve
is $1/(T\alpha)$,
and the area under the curve in the domain, $[\bar q(t), q(T)]$,
is equal to $t/T$,
as shown in Fig. \ref{fig:massq}.

@@

    \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
    \draw[->] (0, 0) -- node[below, inner sep=7mm] {$\bar q$} (2.8, 0);
    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 1.9);

    % xtics
    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
    \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
  \end{tikzpicture}
  %\makebox[\linewidth][c]{
  %  \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
  %}
  \caption{
    \label{fig:massq}
    Geometric characterization of the optimal schedule,
    $\alpha(t)$,
    via the normalized mass distribution, $m(\bar q)$.
    %
    \note{
      The figure was produced by \texttt{tikz}.
      % Gnuplot version by doc/fig/massq.gp
    }%
  }
\end{center}
\end{figure}

We may consider the schedule, $\alpha(t)$,
as a variable transformation of
the mass distribution, $m(\bar q)$.
%much like the Legendre transform.
%
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited in $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
%
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$.
%
A few examples are shown in Table \ref{tab:m_and_a}.
%
For example, in the single-bin scheme,
the mass distribution is a single exponential decay,
and we would have the inverse-time schedule.
%
On the other hand,
if the mass distribution is inversely proportional to $\bar q$,
which is approximately true for the Gaussian updating scheme
at large $\bar q$,
then we would have an exponentially decaying schedule.

\begin{table}[h]\footnotesize
  \caption{\label{tab:m_and_a}
    Variable transformation between the
    reduced schedule, $\alpha^*(t^*)$,
    and the reduced mass distribution, $m^*(\bar q^*)$,
    as defined in Eqs. \eqref{eq:mQ_invTa}
    and \eqref{eq:intmQ_tT}.
  }
  \setlength{\tabcolsep}{3pt} % spacing between columns
  \renewcommand\arraystretch{2.3} % spacing between rows
  \begin{tabular} { c c c }
    \hline
    $m^*(\bar q^*) = -\dfrac{d t^*}{d\bar q^*}$ &
    $\alpha^*(t^*) = -\dfrac{d\bar q^*}{d t^*}$ &
    $\bar q^*(t^*)$
    \\
    \hline
    $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
    $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
    $\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a}$
    %$\,\begin{aligned}
    %  &a, q^*_0 > 0, \\
    %  &\frac{1}{t^*_0} \equiv \left(1+\frac{1}{q^*_0}\right)^a - 1, \\
    %  &c \equiv {q^*_0}^a(t^*_0 + 1)
    %\end{aligned}$
    \\
    $\dfrac{1}{\gamma \, {\bar q}^* + \frac{\gamma}{e^\gamma - 1}}$ &
    $\dfrac{\gamma}{1-e^{-\gamma}} \, e^{-\gamma \, t^*}$ &
    $\dfrac{e^{-\gamma \, t^*} - e^{-\gamma}}{1-e^{-\gamma}}$
    \\
    $\dfrac{\gamma}{1 - e^{-\gamma}} e^{-\gamma \, {\bar q}^*}$ &
    $\dfrac{1}{ \gamma \, t^* + \frac{\gamma}{e^\gamma - 1} }$ &
    $\dfrac{\ln\left[(1-e^{-\gamma}) \, t^* + e^{-\gamma} \right]}{-\gamma}$
    \\
    \hline
    \multicolumn{3}{c}{
    Here,
    $a, q^*_0, \gamma > 0$,
    ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
    and
    $c \equiv {q^*_0}^a(t^*_0 + 1)$.
    } \\
    \hline
  \end{tabular}
\end{table}

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \mathcal Y\bigl( q(T) \bigr)
  &\equiv
  \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(q') \, dq'
  -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  \notag \\
  &
  -\sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  =0
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(q) \, dq
  =
  \frac{ 2 \, C } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

%For simulations with the same $a_0$,
%the shape of the mass distribution, $m(\bar q)$,
%also determines how much the optimal schedule
%would depend on the simulation length, $T$.
%%
%Since both $M(\bar q)$ and $m(\bar q)$ decrease monotonically with $\bar q$,
%Eq. \eqref{eq:opt_qT} shows that
%the optimal $q(T)$ increases with
%the simulation length, $T$,
%more so if $M(\bar q)$ has a long tail.
%%
%Thus, if $M(\bar q)$ decays slowly because of
%many near zero eigenvalues, $\lambda_k$'s,
%as in the metadynamics case,
%both the optimal schedule and
%normalized asymptotic error, $\Err_A(T) \cdot T$,
%would be sensitive to the simulation length, $T$.
%%
%In contrast, if all eigenvalues are $1.0$
%as in the single-bin or WL updating scheme discussed previously,
%the optimal schedule is
%given by the universal inverse-time formula,
%Eq. \eqref{eq:alpha_invt1},
%and the normalized asymptotic error
%is roughly a constant
%[cf. Eq. \eqref{eq:error_singlebin}].
%
%If $M(\bar q)$ decays rapidly,
%the distribution $m(\bar q)$ does not critically
%depend on the cutoff $q(T)$,
%once $q(T)$ is sufficiently large.
%%
%So the optimal schedules of these simulations
%approximately satisfy the scaling relations
%described in Sec. \ref{sec:mass_distr}.
%%
%If, however, $M(\bar q)$ has a long tail
%because of some near-zero $\lambda_k$'s,
%increasing the simulation length $T$
%generally alters the shape of the optimal schedule.

@@

Many commonly-used updating schemes,
such as the Gaussian updating scheme
used in metadynamics,
can be characterized by a rigid window
centered on the current bin.
%
For these translationally invariant or homogeneous
schemes, the eigenvalues can be readily found
from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
as a cosine transform,
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ i = 1 }^b
  \mu_i
  \cos\left(
  \frac{ k \, i \, 2 \, \pi }
       {      g \, n        }
  \right)
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
%
where $g = 1$ here for the periodic variable,
or $2$ for a non-periodic one
(cf. Appendix \ref{sec:more_wband}).

In particular, for the Gaussian updating scheme,
we have\cite{bussi2006}
\begin{align}
  \lambda_k
  &=
  \exp\left(
        -\frac{ k^2 \, \sigma_\varphi^2 }
              {           2             }
      \right)
  =
  \exp\left[
        -\frac 1 2
        \left(
         \frac{ 2 \, \pi \, k \, \sigma_z }
              { g \, \Delta z  }
        \right)^2
      \right]
  ,
%\notag
\label{eq:lambda_Gaussian}
\end{align}
where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
are the width of the Gaussian and span
in the original collective variable, $z$,
respectively.

@@

  \node[draw, text width={width("Integrals, iii")}]
    (Gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};

@@

  \draw[->, >=latex, thick]
    (w) edge (lambda)
    (w) edge (Gamma)
    (sys) edge (Gamma);

  \draw[->, >=latex, thick]
    (lambda) -| (M1)
    (Gamma)  -- (M1)
    (T)      -| (M1)
    (M1)     -- (R1)
    (R1)     |- (qT);

@@

\item \label{step:Gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
(cf. Appendix \ref{sec:Gamma_measure})
as well as the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it can overestimate
  $\epsilon_k$ for large-$k$ modes.
}

\item \label{step:qT}
Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
%
We implemented a variant of the Newton-Raphson method\cite{press3rd}
with the bisection method\cite{press3rd}
as a fallback.\footnote{
The equation was solved for $\ln q(T)$
to avoid negative values of $q(T)$,
and we imposed a lower bound,
$\bigl|\mathcal Y/q(T)\bigr|$, for
$\mathcal Y'\bigl(q(T)\bigr)$
in computing its change.
}
The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
%
%Since $m(\bar q)$ is a decreasing function of $\bar q$,
%we can start from an interval $[Q_L, Q_R]$
%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
%(e.g. we can choose $Q_L = 0$,
%  and repeatedly double $Q_R$ until
%  the latter inequality is satisfied).
%%
%We can then iteratively refine the interval
%by substituting $Q_M = (Q_L + Q_R)/2$
%for either $Q_L$ and $Q_R$,
%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
%until the width of the interval is less than
%the desired precision,
%%
%and set $q(T)$ to $Q_M$.

@@

As shown in Table \ref{tab:lj_error},
both the inverse-time and optimal schedule reduced
the error significantly,
and the errors roughly matched the predicted values
calculated from the precomputed $\Gamma_k$'s and $\epsilon_k$'s.
%
The Gaussian scheme produced less error than
the single-bin scheme in the $\varrho^* = 0.1$ case,
but not in the $\varrho^* = 0.8$ case.
%
In the low density case,
when the local noise contributes significantly to the total error,
the Gaussian scheme
yielded less error than the single bin scheme,
and the optimal schedule was more effective
in reducing the error than the inverse-time schedule.
%
However, in the high-density case
the error was dominated by the $k=1$ eigenmode
corresponding to the longest range motion,
as $\Gamma_1 \approx \Gamma$,
and the reduction of local noise
had little effect on the total error.
%
The bandpass updating scheme behaved similarly,
and in the low density case when
local noise is important,
this scheme (with the mode cutoff $K = 20$)
was superior to the Gaussian counterpart.
